---
title: "Analysis4TwoWayANOVA"
author: "Ellen McMullen"
date: "2025-06-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

Load packages:
```{r}
library(stringr)
library(multcomp)
```

Read in data:
```{r}
all_params = read.csv(
  '../ExpectedResultsAnalysis4/Individuals/CombinedGroups/Models/CombinedGroups_IndividualModels.csv')
```

Split the group names (consisting of the genotype and age labels) into two factors
```{r}
all_params[c('genotype', 'age')] = str_split_fixed(all_params$group, ' ', 2)
```

Two-Way MANOVA (Full Test):
```{r}
# manova
two_way_manova =  manova(cbind(time_activity_parameter_0,
                               time_activity_parameter_1,
                               time_activity_parameter_2,
                               time_coverage_parameter_0,
                               time_coverage_parameter_1,
                               time_p_plus_plus_given_plus_parameter_0,
                               time_p_plus_plus_given_plus_parameter_1,
                               time_p_plus_minus_given_plus_parameter_0,
                               time_p_plus_minus_given_plus_parameter_1,
                               time_p_plus_zero_given_plus_parameter_0,
                               time_p_plus_zero_given_plus_parameter_1,
                               time_p_zero_plus_given_zero_parameter_0,
                               time_p_zero_plus_given_zero_parameter_1,
                               time_p_zero_zero_given_zero_parameter_0,
                               time_p_zero_zero_given_zero_parameter_1,
                               percent_coverage_activity_parameter_0,
                               percent_coverage_activity_parameter_1,
                               percent_coverage_activity_parameter_2,
                               percent_coverage_p_plus_plus_given_plus_parameter_0,
                               percent_coverage_p_plus_plus_given_plus_parameter_1,
                               percent_coverage_p_plus_minus_given_plus_parameter_0,
                               percent_coverage_p_plus_minus_given_plus_parameter_1,
                               percent_coverage_p_plus_zero_given_plus_parameter_0,
                               percent_coverage_p_plus_zero_given_plus_parameter_1,
                               percent_coverage_p_zero_plus_given_zero_parameter_0,
                               percent_coverage_p_zero_plus_given_zero_parameter_1,
                               percent_coverage_p_zero_zero_given_zero_parameter_0,
                               percent_coverage_p_zero_zero_given_zero_parameter_1)
                         ~ genotype*age,
                         data = all_params)
summary(two_way_manova)
```

Now for each relationship of interest, we want to run the MANOVA sub-test:
```{r}
subtest1 =  manova(cbind(time_activity_parameter_0,
                         time_activity_parameter_1,
                         time_activity_parameter_2)
                   ~ genotype*age,
                   data = all_params)
summary(subtest1)
```



Since the MANOVA sub-test is significant, we want to run follow-up ANOVAs on each parameter
```{r}
# time vs activity A
mod1 =  aov(time_activity_parameter_0 ~ genotype*age, data = all_params)
summary(mod1)
# time vs activity B
mod2 =  aov(time_activity_parameter_1 ~ genotype*age, data = all_params)
summary(mod2)
# time vs activity C
mod3 =  aov(time_activity_parameter_2 ~ genotype*age, data = all_params)
summary(mod3)
```
We also want to run a Tukey's HSD test as a post-hoc to detect which groups are different from one another, and the direction of the difference.

```{r}
# time vs activity A post hoc
ph1 = TukeyHSD(aov(time_activity_parameter_0 ~ age+genotype, data = all_params),
               which = "age")
ph1
ph2 = TukeyHSD(aov(time_activity_parameter_0 ~ age+genotype, data = all_params),
               which = "genotype")
ph2
# time vs activity B post hoc
ph3 = TukeyHSD(aov(time_activity_parameter_1 ~ age+genotype, data = all_params),
               which = "age")
ph3
ph4 = TukeyHSD(aov(time_activity_parameter_1 ~ age+genotype, data = all_params),
               which = "genotype")
ph4
# time vs activity C post hoc
ph5 = TukeyHSD(aov(time_activity_parameter_2 ~ age+genotype, data = all_params),
               which = "age")
ph5
ph6 = TukeyHSD(aov(time_activity_parameter_2 ~ age+genotype, data = all_params),
               which = "genotype")
ph6
```


Repeat for other relationships:
```{r}
subtest2 =  manova(cbind(time_coverage_parameter_0,
                         time_coverage_parameter_1)
                   ~ genotype*age,
                   data = all_params)
summary(subtest2)

# time vs coverage A
mod4 =  aov(time_coverage_parameter_0 ~ genotype*age, data = all_params)
summary(mod4)
# time vs coverage B
mod5 =  aov(time_coverage_parameter_1 ~ genotype*age, data = all_params)
summary(mod5)

# time vs coverage A post hoc
ph7 = TukeyHSD(aov(time_coverage_parameter_0 ~ age+genotype, data = all_params),
               which = "age")
ph7
ph8 = TukeyHSD(aov(time_coverage_parameter_0 ~ age+genotype, data = all_params),
               which = "genotype")
ph8
# time vs coverage B post hoc
ph9 = TukeyHSD(aov(time_coverage_parameter_1 ~ age+genotype, data = all_params),
               which = "age")
ph9
ph10 = TukeyHSD(aov(time_coverage_parameter_1 ~ age+genotype, data = all_params),
               which = "genotype")
ph10
```

```{r}
subtest3 =  manova(cbind(time_p_plus_plus_given_plus_parameter_0,
                         time_p_plus_plus_given_plus_parameter_1)
                   ~ genotype*age,
                   data = all_params)
summary(subtest3)

# time vs p++ A
mod6 =  aov(time_p_plus_plus_given_plus_parameter_0 ~ genotype*age,
            data = all_params)
summary(mod6)
# time vs p++ B
mod7 =  aov(time_p_plus_plus_given_plus_parameter_1 ~ genotype*age,
            data = all_params)
summary(mod7)

# time vs p++ A post hoc
ph11 = TukeyHSD(aov(time_p_plus_plus_given_plus_parameter_0 ~ age+genotype,
                    data = all_params), which = "age")
ph11
ph12 = TukeyHSD(aov(time_p_plus_plus_given_plus_parameter_0 ~ age+genotype,
                    data = all_params), which = "genotype")
ph12
# time vs p++ B post hoc
ph13 = TukeyHSD(aov(time_p_plus_plus_given_plus_parameter_1 ~ age+genotype,
                    data = all_params), which = "age")
ph13
ph14 = TukeyHSD(aov(time_p_plus_plus_given_plus_parameter_1 ~ age+genotype,
                    data = all_params), which = "genotype")
ph14
```

```{r}
subtest4 =  manova(cbind(time_p_plus_minus_given_plus_parameter_0,
                         time_p_plus_minus_given_plus_parameter_1)
                   ~ genotype*age,
                   data = all_params)
summary(subtest4)

# time vs p+- A
mod8 =  aov(time_p_plus_minus_given_plus_parameter_0 ~ genotype*age,
            data = all_params)
summary(mod8)
# time vs p+- B
mod9 =  aov(time_p_plus_minus_given_plus_parameter_1 ~ genotype*age,
            data = all_params)
summary(mod9)

# time vs p+- A post hoc
ph15 = TukeyHSD(aov(time_p_plus_minus_given_plus_parameter_0 ~ age+genotype,
                    data = all_params), which = "age")
ph15
ph16 = TukeyHSD(aov(time_p_plus_minus_given_plus_parameter_0 ~ age+genotype,
                    data = all_params), which = "genotype")
ph16
# time vs p+- B post hoc
ph17 = TukeyHSD(aov(time_p_plus_minus_given_plus_parameter_1 ~ age+genotype,
                    data = all_params), which = "age")
ph17
ph18 = TukeyHSD(aov(time_p_plus_minus_given_plus_parameter_1 ~ age+genotype,
                    data = all_params), which = "genotype")
ph18
```

```{r}
subtest5 =  manova(cbind(time_p_plus_zero_given_plus_parameter_0,
                         time_p_plus_zero_given_plus_parameter_1)
                   ~ genotype*age,
                   data = all_params)
summary(subtest5)

# time vs p+0 A
mod10 =  aov(time_p_plus_zero_given_plus_parameter_0 ~ genotype*age,
             data = all_params)
summary(mod10)
# time vs p+0 B
mod11 =  aov(time_p_plus_zero_given_plus_parameter_1 ~ genotype*age,
             data = all_params)
summary(mod11)

# time vs p+0 A post hoc
ph19 = TukeyHSD(aov(time_p_plus_zero_given_plus_parameter_0 ~ age+genotype,
                    data = all_params), which = "age")
ph19
ph20 = TukeyHSD(aov(time_p_plus_zero_given_plus_parameter_0 ~ age+genotype,
                    data = all_params), which = "genotype")
ph20
# time vs p+0 B post hoc
ph21 = TukeyHSD(aov(time_p_plus_zero_given_plus_parameter_1 ~ age+genotype,
                    data = all_params), which = "age")
ph21
ph22 = TukeyHSD(aov(time_p_plus_zero_given_plus_parameter_1 ~ age+genotype,
                    data = all_params), which = "genotype")
ph22
```

```{r}
subtest6 =  manova(cbind(time_p_zero_plus_given_zero_parameter_0,
                         time_p_zero_plus_given_zero_parameter_1)
                   ~ genotype*age,
                   data = all_params)
summary(subtest6)

# time vs p0+ A
mod12 =  aov(time_p_zero_plus_given_zero_parameter_0 ~ genotype*age, 
             data = all_params)
summary(mod12)
# time vs p0+ B
mod13 =  aov(time_p_zero_plus_given_zero_parameter_1 ~ genotype*age, 
             data = all_params)
summary(mod13)

# time vs p0+ A post hoc
ph23 = TukeyHSD(aov(time_p_zero_plus_given_zero_parameter_0 ~ age+genotype, 
                    data = all_params), which = "age")
ph23
ph24 = TukeyHSD(aov(time_p_zero_plus_given_zero_parameter_0 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph24
# time vs p0+ B post hoc
ph25 = TukeyHSD(aov(time_p_zero_plus_given_zero_parameter_1 ~ age+genotype, 
                    data = all_params), which = "age")
ph25
ph26 = TukeyHSD(aov(time_p_zero_plus_given_zero_parameter_1 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph26
```

```{r}
subtest7 =  manova(cbind(time_p_zero_zero_given_zero_parameter_0,
                         time_p_zero_zero_given_zero_parameter_1)
                   ~ genotype*age,
                   data = all_params)
summary(subtest7)

# time vs p00 A
mod14 =  aov(time_p_zero_zero_given_zero_parameter_0 ~ genotype*age, 
             data = all_params)
summary(mod14)
# time vs p00 B
mod15 =  aov(time_p_zero_zero_given_zero_parameter_1 ~ genotype*age, 
             data = all_params)
summary(mod15)

# time vs p00 A post hoc
ph27 = TukeyHSD(aov(time_p_zero_zero_given_zero_parameter_0 ~ age+genotype, 
                    data = all_params), which = "age")
ph27
ph28 = TukeyHSD(aov(time_p_zero_zero_given_zero_parameter_0 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph28
# time vs p00 B post hoc
ph29 = TukeyHSD(aov(time_p_zero_zero_given_zero_parameter_1 ~ age+genotype, 
                    data = all_params), which = "age")
ph29
ph30 = TukeyHSD(aov(time_p_zero_zero_given_zero_parameter_1 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph30
```

```{r}
subtest8 =  manova(cbind(percent_coverage_activity_parameter_0,
                         percent_coverage_activity_parameter_1,
                         percent_coverage_activity_parameter_2)
                   ~ genotype*age,
                   data = all_params)
summary(subtest8)

# percent_coverage vs activity A
mod16 =  aov(percent_coverage_activity_parameter_0 ~ genotype*age, 
             data = all_params)
summary(mod16)
# percent_coverage vs activity B
mod17 =  aov(percent_coverage_activity_parameter_1 ~ genotype*age, 
             data = all_params)
summary(mod17)
# percent_coverage vs activity C
mod18 =  aov(percent_coverage_activity_parameter_2 ~ genotype*age,
             data = all_params)
summary(mod18)

# percent_coverage vs activity A post hoc
ph31 = TukeyHSD(aov(percent_coverage_activity_parameter_0 ~ age+genotype, 
                    data = all_params), which = "age")
ph31
ph32 = TukeyHSD(aov(percent_coverage_activity_parameter_0 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph32
# percent_coverage vs activity B post hoc
ph33 = TukeyHSD(aov(percent_coverage_activity_parameter_1 ~ age+genotype, 
                    data = all_params), which = "age")
ph33
ph34 = TukeyHSD(aov(percent_coverage_activity_parameter_1 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph34
# percent_coverage vs activity C post hoc
ph35 = TukeyHSD(aov(percent_coverage_activity_parameter_2 ~ age+genotype, 
                    data = all_params), which = "age")
ph35
ph36 = TukeyHSD(aov(percent_coverage_activity_parameter_2 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph36
```

```{r}
subtest9 =  manova(cbind(percent_coverage_p_plus_plus_given_plus_parameter_0,
                         percent_coverage_p_plus_plus_given_plus_parameter_1)
                   ~ genotype*age,
                   data = all_params)
summary(subtest9)

# percent_coverage vs p++ A
mod19 =  aov(percent_coverage_p_plus_plus_given_plus_parameter_0 ~ genotype*age, 
             data = all_params)
summary(mod19)
# percent_coverage vs p++ B
mod20 =  aov(percent_coverage_p_plus_plus_given_plus_parameter_1 ~ genotype*age, 
             data = all_params)
summary(mod20)

# percent_coverage vs p++ A post hoc
ph37 = TukeyHSD(aov(percent_coverage_p_plus_plus_given_plus_parameter_0 ~ age+genotype, 
                    data = all_params), which = "age")
ph37
ph38 = TukeyHSD(aov(percent_coverage_p_plus_plus_given_plus_parameter_0 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph38
# percent_coverage vs p++ B post hoc
ph39 = TukeyHSD(aov(percent_coverage_p_plus_plus_given_plus_parameter_1 ~ age+genotype, 
                    data = all_params), which = "age")
ph39
ph40 = TukeyHSD(aov(percent_coverage_p_plus_plus_given_plus_parameter_1 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph40
```

```{r}
subtest10 =  manova(cbind(percent_coverage_p_plus_minus_given_plus_parameter_0,
                          percent_coverage_p_plus_minus_given_plus_parameter_1)
                    ~ genotype*age,
                    data = all_params)
summary(subtest10)

# percent_coverage vs p+- A
mod21 =  aov(percent_coverage_p_plus_minus_given_plus_parameter_0 ~ genotype*age, 
             data = all_params)
summary(mod21)
# percent_coverage vs p+- B
mod22 =  aov(percent_coverage_p_plus_minus_given_plus_parameter_1 ~ genotype*age, 
             data = all_params)
summary(mod22)

# percent_coverage vs p+- A post hoc
ph41 = TukeyHSD(aov(percent_coverage_p_plus_minus_given_plus_parameter_0 ~ age+genotype, 
                    data = all_params), which = "age")
ph41
ph42 = TukeyHSD(aov(percent_coverage_p_plus_minus_given_plus_parameter_0 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph42
# percent_coverage vs p+- B post hoc
ph43 = TukeyHSD(aov(percent_coverage_p_plus_minus_given_plus_parameter_1 ~ age+genotype, 
                    data = all_params), which = "age")
ph43
ph44 = TukeyHSD(aov(percent_coverage_p_plus_minus_given_plus_parameter_1 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph44
```

```{r}
subtest11 =  manova(cbind(percent_coverage_p_plus_zero_given_plus_parameter_0,
                          percent_coverage_p_plus_zero_given_plus_parameter_1)
                    ~ genotype*age,
                    data = all_params)
summary(subtest11)

# percent_coverage vs p+0 A
mod23 =  aov(percent_coverage_p_plus_zero_given_plus_parameter_0 ~ genotype*age, 
             data = all_params)
summary(mod23)
# percent_coverage vs p+0 B
mod24 =  aov(percent_coverage_p_plus_zero_given_plus_parameter_1 ~ genotype*age, 
             data = all_params)
summary(mod24)

# percent_coverage vs p+0 A post hoc
ph45 = TukeyHSD(aov(percent_coverage_p_plus_zero_given_plus_parameter_0 ~ age+genotype, 
                    data = all_params), which = "age")
ph45
ph46 = TukeyHSD(aov(percent_coverage_p_plus_zero_given_plus_parameter_0 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph46
# percent_coverage vs p+0 B post hoc
ph47 = TukeyHSD(aov(percent_coverage_p_plus_zero_given_plus_parameter_1 ~ age+genotype, 
                    data = all_params), which = "age")
ph47
ph48 = TukeyHSD(aov(percent_coverage_p_plus_zero_given_plus_parameter_1 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph48
```

```{r}
subtest12 =  manova(cbind(percent_coverage_p_zero_plus_given_zero_parameter_0,
                          percent_coverage_p_zero_plus_given_zero_parameter_1)
                    ~ genotype*age,
                    data = all_params)
summary(subtest12)

# percent_coverage vs p0+ A
mod25 =  aov(percent_coverage_p_zero_plus_given_zero_parameter_0 ~ genotype*age, 
             data = all_params)
summary(mod25)
# percent_coverage vs p0+ B
mod26 =  aov(percent_coverage_p_zero_plus_given_zero_parameter_1 ~ genotype*age, 
             data = all_params)
summary(mod26)

# percent_coverage vs p0+ A post hoc
ph49 = TukeyHSD(aov(percent_coverage_p_zero_plus_given_zero_parameter_0 ~ age+genotype, 
                    data = all_params), which = "age")
ph49
ph50 = TukeyHSD(aov(percent_coverage_p_zero_plus_given_zero_parameter_0 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph50
# percent_coverage vs p0+ B post hoc
ph51 = TukeyHSD(aov(percent_coverage_p_zero_plus_given_zero_parameter_1 ~ age+genotype, 
                    data = all_params), which = "age")
ph51
ph52 = TukeyHSD(aov(percent_coverage_p_zero_plus_given_zero_parameter_1 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph52
```

```{r}
subtest13 =  manova(cbind(percent_coverage_p_zero_zero_given_zero_parameter_0,
                          percent_coverage_p_zero_zero_given_zero_parameter_1)
                    ~ genotype*age,
                    data = all_params)
summary(subtest13)

# percent_coverage vs p00 A
mod27 =  aov(percent_coverage_p_zero_zero_given_zero_parameter_0 ~ genotype*age, 
             data = all_params)
summary(mod27)
# percent_coverage vs p00 B
mod28 =  aov(percent_coverage_p_zero_zero_given_zero_parameter_1 ~ genotype*age, 
             data = all_params)
summary(mod28)

# percent_coverage vs p00 A post hoc
ph53 = TukeyHSD(aov(percent_coverage_p_zero_zero_given_zero_parameter_0 ~ age+genotype, 
                    data = all_params), which = "age")
ph53
ph54 = TukeyHSD(aov(percent_coverage_p_zero_zero_given_zero_parameter_0 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph54
# percent_coverage vs p00 B post hoc
ph55 = TukeyHSD(aov(percent_coverage_p_zero_zero_given_zero_parameter_1 ~ age+genotype, 
                    data = all_params), which = "age")
ph55
ph56 = TukeyHSD(aov(percent_coverage_p_zero_zero_given_zero_parameter_1 ~ age+genotype, 
                    data = all_params), which = "genotype")
ph56
```

